---
title: "Exploring Discretionary Accruals"
author: "Joachim Gassen"
institute: "TRR 266 Accounting for Transparency"
date: last-modified
date-format: long
fontsize: "9pt"
pdf-engine: xelatex # pdflatex creates rastered fonts
format: 
  beamer:
    slide-level: 3
    number-sections: false
    toc: false
classoption: "aspectratio=169"
execute:
  echo: false
  message: false
  warning: false
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
- \input{beamer_theme_trr266_16x9.sty}
---


```{r setup}
library(tidyverse)
library(glue)
library(gt)

load("../output/results.rda")

convert_to_latex <- function(gt_table) {
	gt_table %>% 
		as_latex() %>% 
		strsplit(split = "\n", fixed= TRUE) %>%
		unlist() %>%
		str_replace(fixed("{tabular*}"), "{tabular}") %>%
		str_replace(fixed("{\\linewidth}{@{\\extracolsep{\\fill}}"), "{")
}
```


# Motivation

### Discretionary Accruals

- Are used in tons of papers
- As measurement constructs for almost everything
  - Earnings Management
  - Earnings Quality
  - Audit Quality
  - Management Quality
- But how do they look like?


### Data

```{r tab_descriptives_sample, results="asis"}
latex_tab <- c(
	"\\centering{",
  head(convert_to_latex(tab_desc_stat)[-(1:2)], -1),
	"}"
)
cat(paste(latex_tab, collapse = "\n"))
```


### Distribution of Discretionary Accruals over time - unrestricted sample

```{r BPlotFULL, fig.height=5}
fig_boxplot_full
```


### Distribution of Discretionary Accruals over time - analysis sample

```{r BPlotSMP, fig.height=5}
fig_boxplot_smp
```


### Correlation of Modified Jones with Dechow and Dichev DA

```{r CorrVizz, fig.height=5}
fig_scatter_md_dd
```


### Association of Dechow and Dichev DA with firm size

```{r AssDDSize, fig.height=5}
fig_scatter_dd_lnta
```


### No presentation without correlation table...

``` {r CorrTable, results="asis"}
latex_tab <- c(
	"\\centering{",
  head(convert_to_latex(tab_corr)[-(1:2)], -1),
	"}"
)
cat(paste(latex_tab, collapse = "\n"))
```


### ... and regression table

``` {r RegTable, results="asis"}
tab_reg <- head(convert_to_latex(tab_regression)[-(1:2)], -1)

#  Add a midline above the Observations row
obs_row <- which(str_detect(tab_reg, fixed("Observations")))
tab_reg <- c(
	tab_reg[1:(obs_row-1)], "\\midrule", tab_reg[obs_row:length(tab_reg)]
)

latex_tab <- c(
	"\\centering{",
  tab_reg,
	"}"
)
cat(paste(latex_tab, collapse = "\n"))
```


### Conclusion

- Discretionary accruals are simply wonderful! 
- With these distributional properties and the power of large samples they will almost always be significant in one direction or the other regardless the setting or controls
- Use them for every economic construct under the sun!







